% Complete documentation on the extended LaTeX markup used for Python
% documentation is available in ``Documenting Python'', which is part
% of the standard documentation for Python.  It may be found online
% at:
%
%     http://www.python.org/doc/current/doc/doc.html

\documentclass{manual}
\RequirePackage[latin9]{inputenc}
\usepackage{graphicx}

\title{AlphaFlow manual}

% Please at least include a long-lived email address;
% the rest is at your discretion.
\authoraddress{
    gocept GmbH \& Co. KG\\
    Forsterstraﬂe 29\\
    06112 Halle (Saale)\\
    E-Mail: \email{ct@gocept.com}
}

\date{\today}     % update before release!
                % Use an explicit date so that reformatting
                % doesn't cause a new date to be used.  Setting
                % the date to \today can be used during draft
                % stages to make it easier to handle versions.

\release{2.0.0}           % release version; this is used to define the
                          % \version macro

\makeindex          % tell \index to actually write the .idx file

\begin{document}

\maketitle

% This makes the contents more accessible from the front page of the HTML.
\ifhtml
\chapter*{Front matter\label{front}}
\fi

\copyright 2004-2006 gocept gmbh \& co. kg

Licensed under the Zope Public License, Version 2.1 (the ``License''); you may not use this file except in compliance with the License. You may obtain a copy of the License at \url{http://www.zope.org/Resources/ZPL}.

THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS ``AS IS'' AND ANY EXPRESSED
OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE IMPLIED WARRANTIES OF
MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO
EVENT SHALL THE COPYRIGHT HOLDERS BE LIABLE FOR ANY DIRECT, INDIRECT,
INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT
LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR SERVICES; LOSS OF USE, DATA, OR
PROFITS; OR BUSINESS INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF
LIABILITY, WHETHER IN CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE
OR OTHERWISE) ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF
ADVISED OF THE POSSIBILITY OF SUCH DAMAGE.

\begin{abstract}

\noindent

AlphaFlow is a simple, flexible and easy-to-customize workflow engine for
Plone. It implements an activity based workflow concept to manage content
objects in Plone.

This document is the official documentation for developers and administrators
using AlphaFlow. It contains a tutorial, a user guide and a reference.

\begin{seealso}
    \seetitle[http://www.gocept.com/open\_source\_software/AlphaFlow/]{AlphaFlow homepage}{for up to date news, releases and information about AlphaFlow.}
\end{seealso}
\end{abstract}

\tableofcontents


\include{introduction}

\include{installation}

\include{tutorial}

\include{architecture}

\include{admin}

% Activity reference
\include{activities}

% API reference
\include{api}

\include{recipes}

\chapter{Reporting bugs}

If you find a problem in AlphaFlow, we would love to fix it. You can use the
mailinglist to discuss whether the problem is a bug or not and how it should be
fixed. But to allow us to track problems and react appropriately, please file
any problem you find in the gocept bugtracker (\url{http://bugs.gocept.com}). 

\begin{itemize}

    \item If you submit a bug in our tracker for the first time, please make
        sure you create an account with your email address as username.

    \item Then use the ``search'' action to check if the problem you found has
        already been submitted by someone else. Please do not file duplicate
        bugs, but comment existing bugs and tell us about your experience with
        the problem.

    \item If the problem is not reported yet, please click the ``New'' action to create a new bug entry.

    \item Then select ``AlphaFlow'' as the relevant product to file the bug into.

    \item In the following form, you have to fill out a couple of fields that are mandatory:
        \begin{description}
        \item[Version] where you indicate which version of AlphaFlow you use
        \item[Summary] should give a short and distinguishable description of the problem in (less than) one sentence
        \item[Description] As much relevant information about the problem as you have. Typically this might include code examples, tracebacks and other
            textual explanations. Try to describe the problem as detailed as possible to allow us to reproduce it.
        \end{description}

        The form also has a couple of links that will explain the other fields
        to you, allowing even better specified problem reports.

    \item At last, click on the button ``Commit'', and you have successfully reported the problem to us.
\end{itemize}

Once you have filed a problem in the bug tracker, you will automatically get
informed when we work on the problem and when we fix it. We also use it to
communicate with you to get more information or tell you about possible short
term fixes.

\chapter{Getting support}

\begin{description}
   \item[Mailinglist]
       Send an email with the word ``subscribe'' in the subject to 
       \email{alphaflow@lists.gocept.com} to join the AlphaFlow mailinglist.
       
   \item[Commercial support]
        Commercial support is available from gocept
        (\url{http://www.gocept.com}) in various flavours. Please email us at
        support@gocept.com or call +49 345 122 98 89 0 to get an offer for
        training, consulting, support contracts or custom developments.
\end{description}


\end{document}
